Social inequality in mortality among adults and elderly in northern Sweden 1851--2013\footnote{'Socoeconomic inequalities and mortality I: Europe'. 41st Annual Meeting of the Social Science History Association, Chicago, November 17--20, 2016.}

### 2017-05-18 oldmort.Rnw
refresh

<<readtab>>=
load("data/vbTab.rda") 
vbTab$period <- factor(vbTab$period)
levels(vbTab$hisclass40) <- c("elite", "lowMan", "lowWhiteC", "worker")
levels(vbTab$hisclass60) <- c("elite", "lowMan", "lowWhiteC", "worker")
levels(vbTab$period)[8] <- "1991-2001"
load("data/vb.rda")
vb$period <- factor(vb$period)
levels(vb$hisclass) <- c("elite", "lowMan", "lowWhiteC", "worker")
levels(vb$hisclass40) <- c("elite", "lowMan", "lowWhiteC", "worker")
levels(vb$hisclass60) <- c("elite", "lowMan", "lowWhiteC", "worker")
levels(vb$period)[8] <- "1991-2001"
@

@@

<<hiscskum, fig.cap = "Exposure by hisclass (from the top: elite, farmers and lower business, lower white collar, workers) and time period", fig.height = 4.5>>=
xx <- with(vbTab, tapply(exposure, list(period = period, hisclass = hisclass40), sum))
yy <- round(prop.table(xx, 1) * 100, 1)
barplot(t(yy)[4:1, ], col = c("red", "blue", "green", "black"), las = 2, ylab = "Per cent")
@

@@

<<nana, results = 'asis'>>=
library(xtable)
load("data/vb0.rda")
vb0$hisclass <- factor(vb0$hisclass, labels = c("elite", "lowMan", "lowWhiteC", "worker"))
x <- with(vb0, table(period, hisclass, useNA = "ifany"))
x <- round(prop.table(x, 1) * 100)
xtable(x, caption = "Hisclass by time period, per cent.", label = "tab:nana", digits = 0)
@

@@

<<colorc>>=
## Colors for plotting strata:
scol <- c("blue", rep("darkgreen", 2), "red")
## Common for all groups:

@@

<<cumsw, fig.cap = cap0, fig.scap = scap0, fig.height = 10>>=
source("R/plotED.glm.R")
oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, ylab = "H(65|40)",
                main = "Women")##, exclude = 5:6)
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, ylab = "H(65|40)",
                main = "Men")##, exclude = 5:6)
par(oldpar)
@

>>>>

<<readtab>>=
load("data/vbTab.rda") 
hisclevels <-  c("elite", "middle", "worker") # NOTE!!
vbTab$period <- factor(vbTab$period)
levels(vbTab$hisclass40) <- hisclevels
levels(vbTab$hisclass60) <- hisclevels
levels(vbTab$period)[8] <- "1991-2001"
load("data/vb.rda")
vb$period <- factor(vb$period)
levels(vb$hisclass) <- hisclevels
levels(vb$hisclass40) <- hisclevels
levels(vb$hisclass60) <- hisclevels
levels(vb$period)[8] <- "1991-2001"
@

@@

<<hiscskum, fig.cap = "Exposure by hisclass (from the top: elite, farmers and lower business, lower white collar, workers) and time period", fig.height = 4.5>>=
xx <- with(vbTab, tapply(exposure, list(period = period, hisclass = hisclass40), sum))
yy <- round(prop.table(xx, 1) * 100, 1)
who <- length(levels(vbTab$hisclass40))
barplot(t(yy)[who:1, ], col = c("red", "blue", "black"), las = 2, ylab = "Per cent")
@

@@

<<nana, results = 'asis'>>=
library(xtable)
load("data/vb0.rda")
vb0$hisclass <- factor(vb0$hisclass, labels = hisclevels)
x <- with(vb0, table(period, hisclass, useNA = "ifany"))
x <- round(prop.table(x, 1) * 100)
xtable(x, caption = "Hisclass by time period, per cent.", label = "tab:nana", digits = 0)
@

@@

<<colorc>>=
## Colors for plotting strata:
scol <- c("black", "blue", "red")
## Common for all groups:

@@

<<cumsw, fig.cap = cap0, fig.scap = scap0, fig.height = 10>>=
source("R/plotED.glm.R")
oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = TRUE, ylab = "log(1-S(65|40))",
                main = "Women 40-64")##, exclude = 5:6)
moutp <- plotED.glm(resm, col = scol, logScale = TRUE, ylab = "log(1-S(65|40))",
                main = "Men 40-64")##, exclude = 5:6)
par(oldpar)
@
